% !TeX root = heatdiscrete.tex
\section{Randomized computational models}
\label{sec:applications}
In this section we show the connection between
\autoref{thm:blakley-dixon}, \autoref{thm:main} and 
the randomized communication and query complexities 
of the Hamming distance problem.

\subsection{Communication complexity}
In a two player communication problem the players, 
named Alice and Bob, receive separate inputs, 
respectively $x\in \mathcal{X}$ and $y\in \mathcal{Y}$, 
and they communicate in order to compute the value 
$f(x,y)$ of a function
$f\colon \mathcal{X}\times \mathcal{Y}\to\binary$ 
(known to both players). 
In an $r$-round protocol, the players can take at most 
$r$ turns alternately sending each other a message 
(that is, a bit string) and the last player to receive 
a message declares the output of the protocol.
A protocol can be {\em deterministic} or {\em randomized}; 
in the latter case the players can base their actions on 
a common random source and we measure the 
{\em error probability}: the maximum over inputs 
$(x,y)\in \mathcal{X}\times\mathcal{Y}$, of the probability 
that the output of the protocol differs from $f(x,y)$. 
The {\em communication cost} of a protocol is the maximum, 
over the inputs and the random string, of the total number 
of bits sent between the players.
For a function 
$f\colon \mathcal{X}\times\mathcal{Y}\to\binary$, 
an integer $r$ and $\delta\in[0,1]$, we denote by 
$R^r_{\delta}(f)$ the minimum over all protocols
for $f$ having $r$-rounds and error probability at most 
$\delta$, of the communication cost incurred. We define 
$R_{\delta}(f)$ similarly, but we take the maximum over 
$\delta$-error protocols with no restriction on the number 
of rounds it uses.

In the $k$-Hamming distance problem, denoted $\Ham^n_k$,
the players receive length-$n$ bit strings, respectively 
$x,y\in\cube$, and are required determine if 
$\lone{x-y}\le k$ or not.
There is a well known one-round communication protocol
which accomplishes this with error probability $\delta$ 
by communicating $O(k\log\nparen{k/\delta})$ bits.

\begin{theorem}
[e.g., Huang, Shi, Zhang and Zhu \cite{HuangSZZ2006}]
\label{thm:ub}
It holds that $$R^1_\delta(\Ham^n_k) = 
O(\min\set{k\log\nparen{k/\delta}, k\log (n/k)}).$$
\end{theorem}

Highly related to the $\Ham^n_k$ is the $k$-disjointness 
problem $\Disj^n_k$, wherein the players each receive a 
$k$-subset of $[n]$ and their goal is to determine if their 
sets intersect.
Notice that $\Disj^n_k$ can be seen as a promise version of
$\Ham^n_{2k-2}$ where each player is guaranteed to have a string
with Hamming weight $k$: the sets are disjoint if and only if the
Hamming distance between the characteristic vectors of the sets
is more than $2k-2$. Therefore any upper bound for the $\Ham^n_k$ 
carries over to $\Disj^n_k$ and any lower bound for $\Disj^n_k$
carries over to $\Ham^n_k$.
Around 1993, 
Håstad and Wigderson \cite{HastadW2007} showed that
there is a more efficient protocol for $\Disj^n_k$ than that implied 
by \autoref{thm:ub}, which communicates only $O(k)$ bits, but over
$O(\log k)$ rounds.

On the lower bounds side, the result of Kalyanasundaram and
Schnitger \cite{KalyanasundaramS1992} implies that $\Omega(k)$
bits is needed for these problems even if one uses arbitrarily
large number of round protocols. In \cite{BuhrmanGMW2012} 
Buhrman, Garcia-Soriano, Matsliah and de Wolf have shown
that any 1-round protocol for $\Disj^n_k$ needs to
communicate at least $\Omega(k \log k)$ bits when $k^2<n$ 
(this result was proven later in \cite{DasguptaKS12} also). 
In Theorem~3.2 of \cite{Saglam2011}, an $\Omega(k\log(1/\delta))$
bound for $1$-round complexity of $\Disj^n_k$ was shown even
when Bob receives just one element (i.e., the indexing problem)
for $k<\delta n$ and a slightly more general result was shown in
\cite{JayramW2011}. Finally in \cite{SaglamT2013} the
communication complexity of $\Disj^n_k$ was settled:
\begin{align*}
R^r_{1/3}(\Disj^n_k)=\Theta(k\log^{(r)}k)
\end{align*}
for $1\le r \le \log^*k$ and $k^2<n$.
Their upper bound solves the disjointness problem with 
error probability at most $1/\exp k + 1/\exp^{(r)}(c\log^{(r)}k)$ 
for any $c>1$ by communicating
$O(k \log^{(r)} k)$ bits over $r$ rounds. In fact bulk of the bits
is sent in the first round and the rest of the rounds
amount to an $O(k)$ bits of communication. Taking $r=\log^* k$,
this leads to an $O(k)$ bits protocol with error probability 
that is exponentially small in $k$.
Their lower bound shows that at least one
message of size 
$\Omega(k\log^{(r)} k)$ bits needs to be sent by any 
$r$-round protocol, even if it has error probability $1/3$.
Prior to this work, this lower bound provided the strongest lower
bound for $\Ham^n_k$ also, along with the incomparable bound of 
$\Omega(k\log (1/\delta))$ due to \cite{BlaisBG2014}
which holds for any number of rounds, which we discuss shortly.

\input{hamvsdisj}

To summarize the above results, the 1-round communication 
complexity of both $\Disj^n_k$ and $\Ham^n_k$ is 
$\Theta(k\log(k/\delta))$ by 
\cite{BuhrmanGMW2012, Saglam2011, JayramW2011} and 
\cite{HuangSZZ2006}. We know that $\Disj^n_k$ can be solved 
much more efficiently if one is allowed multiple rounds:
firstly the $\log k$ factor can be removed \cite{HastadW2007} 
and secondly the error probability can be brought down to 
$\exp(-k)$, by using no more than $\log^* k$ rounds 
\cite{SaglamT2013}. It is an interesting question whether 
similar efficiency improvements can be obtained for $\Ham^n_k$ 
also, by using multiple rounds.
The first separation of $\Disj^n_k$ and $\Ham^n_k$ was proven
in \cite{BlaisBG2014}, which shows that $\Omega(k\log(1/\delta))$
lower bound holds for any protocol solving $\Ham^n_k$. 
Therefore in $\Ham^n_k$, we get no improvements in error probability
by interactive communication. It remained an open 
question whether {\em any} improvement can be made 
at all to the 1-round protocol by communicating interactively.
In this work we answer this question negatively: 

\begingroup
\def\thetheorem{\ref{thm:klogk}}
\begin{theorem}[restated]
For $k^2<\delta n$ we have 
$R_{\delta}(\Ham^n_k) = \Omega(k\log (k/\delta))$.
The bound applies even to protocols that may
output an arbitrary answer when $\lone{x-y} \notin \set{k-2, k,k+2}$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
Before we proceed with proving \autoref{thm:klogk}, let us
first warm up by showing that \autoref{thm:blakley-dixon} 
implies an  $\Omega(k\log(1/\delta))$ lower bound on 
$R_\delta(\Ham^n_k)$.
To do so, let us review the so called {\em corruption bound}
method. Let  $f\colon\mathcal{X}\times\mathcal{Y}\to\binary$ 
be the function the players would like to compute with 
Alice having received $x\in \mathcal{X}$ and Bob $y\in \mathcal{Y}$. 
For a protocol $P$ for $f$, define the matrix 
$A_P\colon\mathcal{X}\times\mathcal{Y}\to[0,1]$ 
such that $A_P(x,y)$ is the probability
that the protocol outputs 1 on input $(x,y)$.
It is well known and not difficult to see that
if $P$ has communication cost $c$, then $A_P$ is the
average of matrices each of which is the sum of at most
$2^c$ rank 1 matrices $u v^\transpose$ with 
$u\in \binary^\mathcal{X}$ and $v\in\binary^\mathcal{Y}$.
Therefore to show the communication cost of a protocol
$P$ is more than $c$, it suffices to argue $A_P$ lies
outside $2^c$ times the polytope
\begin{align*}
\mathcal{T}\defeq 
\conv\set{uv^\transpose\mid 
u\in\binary^\mathcal{X}, v\in\binary^\mathcal{Y}},
\end{align*}
where $\conv$ denotes the convex hull. By convexity, 
$A_p$ lies outside of $2^c\mathcal{T}$ if and only if there 
is a hyperplane (with normal $H$) separating the two; 
namely that $\iprod{A_P}{H}>2^c\iprod{R}{H}$ for all 
vertices $R$ of the polytope $\mathcal{T}$.

Let $\mu_k\colon\cube\times\cube\to\realspos$ be the distribution
on pairs $(x,y)$ obtained as follows. Sample $x$ uniformly at random
and obtain $y$ by flipping $k$ coordinates of $x$ chosen uniformly 
at random and with replacement (here if a coordinate gets flipped 
twice it reverts back to its initial value).
\begingroup
\def\thetheorem{\ref{thm:klogdelta}}
\begin{theorem}[restated]
For $k^2<\delta n$ we have 
$R_{\delta}(\Ham^n_k) = \Omega(k\log (1/\delta))$.
The bound applies even to protocols that may
output an arbitrary answer when $\lone{x-y} \notin \set{k,k+2}$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
Suppose we have a randomized protocol for $\Ham^n_k$ with 
error probability $\delta$. Form the matrix $A$, where
$A(x,y)$ is the probability that the protocol reports
$\lone{x-y}\le k$ on input $(x,y)$.

Set $H = \mu_k - \mu_{k+2}/(3\delta)$. Let us first argue that
$\iprod{A}{H}\ge 1/3$. Note that when we sample $k$ elements 
from $[n]$ uniformly at random, by a union bound, the probability
that there is a collision is at most $\binom{k}{2}/n$, therefore 
$\mu_k$ chooses a pair $(x,y)$ at distance $k$ with probability 
at least $1-\binom{k}{2}/n$.
Hence,
$\iprod{A}{\mu_k} 
   \ge (1-\delta)(1-\binom{k}{2}/n)
     > 1-3\delta/2$,
where in the last step we used $k^2< \delta n$.
Similarly, we have $\iprod{A}{\mu_{k+2}} \le 
                    \delta + \binom{k+2}{2}/n \le 3\delta/2$,
thus it follows that $\iprod{A}{H}\ge 1/3$ for $\delta \le 1/9$.

Next we argue that $\iprod{R}{H}< (3\delta)^{k/2}$ for any 
$R=uv^\transpose$ with $u,v\in\cube$. If 
$\iprod{R}{\mu_k}<(3\delta)^{k/2}$, we are done as 
$\iprod{R}{\mu_{k+2}}\ge 0$ and is a negative term in 
$\iprod{R}{H}$. If $\iprod{R}{\mu_k}\ge (3\delta)^{k/2}$
on the other hand,
observing $\iprod{R}{\mu_k}=\iprod{v}{W^k u}/2^n$,
where $W$ is the normalized adjacency matrix of the Hamming cube,
we have by \autoref{thm:blakley-dixon}
\begin{align*}
\nparen{\frac{\lnorm{u}{2}\lnorm{v}{2}}{2^n}}^{2/k}
\iprod{v}{W^{k+2}u} \ge 
\iprod{v}{W^{k}u}^{1+2/k}.
\end{align*}
Note $\lnorm{u}{2}\lnorm{v}{2}\le 2^n$ since $u,v$ are 0-1 vectors,
therefore 
$\iprod{R}{\mu_{k+2}}\ge 3\delta\iprod{R}{\mu_{k}}$
and hence $\iprod{R}{H}\le 0$. In either case we have shown
that $\iprod{R}{H} < (3\delta)^{k/2}$. This implies an 
$\log((3\delta)^{-k/2} / 3) = \Omega(k\log(1/\delta))$ bits 
lower bound on $R_\delta(\Ham^n_k)$.
\end{proof}
Interestingly, \autoref{thm:klogk} cannot be proved by a direct 
application of the corruption method described above. 
If we assume that the protocol
is supposed to output 1 on inputs $\lone{x-y}\le k$, then there are 
vertices of the polytope $\mathcal{T}$ for which 
the $\Omega(k\log(1/\delta))$ bound of 
\autoref{thm:klogdelta} is tight. If we assume that the protocol
is supposed to output 1 on inputs $\lone{x-y}> k$ on the other
hand, no bound above $\Omega(k)$ can be obtained, as there are
vertices for which this is tight.
If we insist however that the protocol outputs 1 for $\lone{x-y}=k$
and 0 for $\lone{x-y}\in\set{k-2,k+2}$ then a protocol with cost
smaller than $O(k\log (k/\delta))$
would be in violation of the near log-convexity principle we 
established in \autoref{thm:main} as we argue next.
Of course, if we had a $\delta$-error randomized protocol 
$P$ for 
$\Ham^{n+2}_k$ outputting 1 when $\lone{x-y}\in\set{k-2,k}$ 
and 0 if $\lone{x-y}=k+2$ 
(but without any guarantees for other types of inputs), then 
given inputs $a,b\in\cube$ Alice and Bob can run $P$
(say, in parallel)
on instances $(00a, 00b)$ and $(00a, 11b)$ and declare 
$\lone{a-b}=k$ if $P$ returns 1 on $(00a, 00b)$ and 
0 on $(00a, 11b)$.
This would lead to a protocol with twice the error probability
and communication cost of $P$, deciding between $\lone{a-b}=k$,
$\lone{a-b}=k-2$ and $\lone{a-b}=k+2$. The table below shows that
$P$ outputting 1 on $(00a, 00b)$ and 0 on $(00a, 11b)$ implies
$\lone{a-b}=k$ or at least one invocation of $P$ erred.

\begin{center}
\def\arraystretch{1.2}
\begin{tabular}{rccc}
Input       & $k-2$ & $k$  & $k+2$ \\\hline
$(00a,00b)$ &   1   &  1   &  0  \\
$(00a,11b)$ &   1   &  0   &  ? 
\end{tabular}
\end{center}

\begin{proof}[Proof of \autoref{thm:klogk}]
Suppose we have a $\delta$-error randomized protocol that 
outputs 1 when $\lone{x-y} = k$ and 0 when 
$\lone{x-y} = k-2$ or $\lone{x-y} = k+2$.
 
Form the matrix $A$, where $A(x,y)$ is the probability 
that the protocol reports that $\lone{x-y}= k$ on 
input $(x,y)$. Let $\alpha_1,\alpha_2> 0$ be some reals so that
\autoref{thm:main} implies
$m_{t+2}\ge t^{\alpha_1}m^{1+2/t}$ or 
$m_{t+2}m_{t-2}\ge \alpha_2{m_t}^2$ for $m_t$ 
defined in statement of this theorem. 

Set $H = \mu_{k} - (\mu_{k-2} + \mu_{k+2})/(6\delta)$.
Let us argue first that $\iprod{A}{H}\ge 1/3$.
One can verify that 
$\iprod{A}{\mu_k}\ge 
  (1-\delta)(1-\binom{k}{2}/n)> 1-3\delta/2$
and $\iprod{A}{\mu_{k-2}} + \iprod{A}{\mu_{k-2}} < 3\delta$.
Hence $\iprod{A}{H}\ge 1/3$ for $\delta \le 1/9$.

We upper bound $\iprod{R}{H}$ for some rank-1
matrix $R=uv^\transpose$ with 0-1 values. Let $W$ be
the normalized adjacency matrix of the Hamming cube graph.
Observe that $\iprod{R}{\mu_k} = \iprod{v}{W^ku}/2^n$.
By \autoref{thm:main}, either 
$\iprod{R}{\mu_{k+2}}\iprod{R}{\mu_{k-2}}
\ge \alpha_2 \iprod{R}{\mu_k}^2$ or 
\begin{align*}
\nparen{\frac{\lnorm{u}{2}\lnorm{v}{2}}{2^n}}^{2/k}
\iprod{R}{\mu_{k+2}}\ge k^{\alpha_1}\iprod{R}{\mu_{k}}^{1+2/k}.
\end{align*}
In the former case,
\begin{align*}
\frac{\iprod{R}{\mu_{k+2}}+\iprod{R}{\mu_{k-2}}}{2}\ge 
\sqrt{\iprod{R}{\mu_{k+2}}\iprod{R}{\mu_{k-2}}}\ge
\sqrt{\alpha_2}\iprod{R}{\mu_k},
\end{align*}
which implies that $\iprod{R}{H}<0$ whenever 
$\delta<2\sqrt{\alpha_2}/6$ (recall $\alpha_2$ is a constant).
In the latter case we get $\iprod{R}{H}<0$ unless
$6\delta\iprod{R}{\mu_{k+2}} \le \iprod{R}{\mu_{k}}$, 
which implies, 
recalling $\lnorm{v}{2}\lnorm{u}{2}\le 2^n$,
that $k^{\alpha_1}\iprod{R}{\mu_k}^{2/k}<6\delta$.
From this we get
\begin{align*}
  \iprod{R}{H}\le \iprod{R}{\mu_k}\le
  \nparen{\frac{6\delta}{k^{\alpha_1}}}^{k/2},
\end{align*}
and hence $\iprod{R}{H} < \nparen{\frac{6\delta}{k^{\alpha_1}}}^{k/2}$ 
in every case and 
$R_\delta(\Ham^n_k)=\Omega(k\log(k/\delta))$
whenever $k^2<\delta n$.
\end{proof}

For a protocol $P$, denote by $\Pi=\Pi(x,y)$ the random variable 
entailing all the messages communicated between the players 
on input $(x,y)$.
So far we have considered the communication cost of a protocol
which is the maximum length of $\Pi$ over all inputs and the 
configurations of the random source (these together determine 
the value of $\Pi$). When a distribution $\mu$
on the inputs is available, we may speak of
a more refined notion of cost, {\em the internal information cost}, 
for a protocol $P$ which is defined as
\begin{align*}
\mathsf{IC}_\mu(P)\defeq \muti{\Pi}{Y\emid X} + \muti{\Pi}{X\emid Y},
\end{align*}
where $(X,Y)\sim\mu$. Combining our 
\autoref{thm:klogk} with a result of \cite{KerenidisLLRX2015}
which relates information and communication costs of a protocol 
under suitable circumstances, one
can conclude that any randomized protocol for $\Ham^n_k$ has
information cost $\Omega(k\log k)$ as well, under the distribution
$\mu = (\mu_k+\mu_{k-2}+\mu_{k+2})/3$. However we note that instead 
of using \autoref{thm:main} black-box,
taking a closer look at the proof of \autoref{thm:blakley-dixon} 
and not performing the relaxation provided in \autoref{lem:negterms},
we get the following more directly.
\begin{theorem}
\label{thm:infocost}
Let $P$ be a protocol outputting 1 on pairs $(x,y)$ having
$\lone{x-y}=k$ with 
probability $1-\delta$ and outputting 0 on pairs 
$(x,y)$ having
$\lone{x-y}\in\set{k-2,k+2}$ with probability $1-\delta$. 
We have 
$\mathsf{IC}_{\mu_k}(P)=\Omega(k\log (k/\delta))$.
\end{theorem}

Let us finally mention another highly related problem, 
the so called the gap Hamming distance problem. 
In $\Ghd^n_{k}$, each of the players receive a bit string,
respectively $x,y\in\cube$, with the promise that either 
$\lone{x-y}\le k$ or $\lone{x-y}\ge k + \sqrt{k}$.
Their goal is to determine which is the case for any given input.
In \cite{ChakrabartiR2012}, an $\Omega(k)$ lower bound
for this problem was shown, which applies to protocols with any
number of rounds.
Here we conjecture an improvement to this bound
and argue that it would follow from a natural analogue of
\autoref{thm:main} for continuous time Markov chains, which
we discuss in \autoref{sec:discussion}.
\begin{conjecture}
\label{conj:ghd}
For $k<\delta n$, we have $R_\delta(\Ghd^n_k) 
= \Omega(k \log(1/\delta))$.
\end{conjecture}

\subsection{Parity decision trees}
\def\PD{\mathrm{PD}}
\def\PS{\mathrm{PS}}
In the parity decision tree model, we are given a string 
$x\in\field_2^n$ and our goal is to determine whether 
$x$ satisfies a fixed predicate
$P\colon\field_2^n\to\binary$ by only making linear 
measurements of the form $\iprod{x}{y}$ for some 
$y\in \field_2^n$ we get to choose. Here, the inner product
is over $\field_2^n$, and therefore we get a single bit 
answer for every measurement we make.

Such measurements can be identified by binary decision
trees wherein each internal node is labeled by a
$y\in\field_2^n$ denoting the linear measurement
$\iprod{x}{y}$ we would make at that node and each leaf
is labeled by a YES or a NO denoting the final decision we 
arrive. Given such a tree and an $x$, the output of the 
decision tree is obtained by a root to leaf walk, where at 
each internal node $v$ with label $y_v$, 
we perform the measurement 
$\iprod{x}{y_v}$ and walk to the left child of $v$ if 
$\iprod{x}{y_v}=0$ and to its right child if 
$\iprod{x}{y_v}=1$. If a leaf node is reached, 
the label of the node is taken as the answer of 
the decision tree. Two quantities we are concerned with
are the depth and the size (i.e., the total number of nodes)
of the tree.

A $\delta$-error randomized decision tree is a distribution
$\nu$ over decision trees such that for any fixed $x$, 
the sampled decision tree outputs the correct answer with 
probability at least $1-\delta$, where the randomness is over 
the choice of the decision tree from $\nu$. 
The depth and the size of a randomized decision tree can be 
taken as the maximum over the decision trees in the support 
of $\nu$ (here, one can also take the average depth or size also;
our result on decision tree size actually lower bounds this 
potentially smaller quantity).

For a predicate $P\colon\field_2^n\to\binary$, let 
$\PD_\delta(P)$ be the minimum, over all randomized decision 
trees $T$ computing $P$ with probability $1-\delta$, of the 
depth of $T$.
Let $\PS_\delta(P)$ be the minimum, over all randomized decision 
trees $T$ computing $P$ with probability $1-\delta$, of the 
size of $T$. The following inequalities are immediate
\begin{align}
  R_\delta(P\circ\oplus)&\le 2\PD_\delta(P)\label{eq:rvspd},\\
  \log\PS_\delta(P)&\le \PD_\delta(P)\nonumber,
\end{align}
where $P\circ \oplus$ is the two player communication game 
in which the two players are given strings $x,y\in\field_2^n$ 
and are required to calculate $P(x+y)$. 
We remark that $\log \PS_\delta$ is incomparable to $R_\delta$ 
in general.

Here we study the predicate $H^n_k$ which equals 1 if 
and only if the Hamming weight of its input is precisely $k$. 
By \autoref{eq:rvspd} and a padding argument similar to the 
one we gave before the proof of \autoref{thm:klogk}, each lower 
bound for $\Ham^n_k$ listed in \autoref{table:hamvsdisj} applies 
to $\PD_\delta(H^n_k)$ as well. In \cite{BlaisK2012} another 
direct $\Omega(k)$ bound for $\PD_\delta(H^n_k)$ was shown. In 
\cite{BhrushundiCK2014}, showing an $\Omega(k\log k)$ lower bound 
to a variant of $\PD_\delta(H^n_k)$ to obtain tight bounds for 
$k$-linearity problem (see \autoref{sec:proptest}) was suggested. 
Finally, our \autoref{thm:klogk} shows that 
$\PD_\delta(H^n_k)=\Omega(k\log(k/\delta))$, which is tight. 
Next we show the same bound holds even for $\log \PS_\delta(H^n_k)$.

\begingroup
\def\thetheorem{\ref{thm:paritysize}}
\begin{theorem}[restated]
For $k^2<\delta n$, 
$\log \PS_\delta(H^n_k)=\Omega(k\log(k/\delta))$.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup
\begin{proof}
\def\f{\field_2^n}
The proof is very similar to that of \autoref{thm:klogk}, so we only 
describe the differences.

Let $T$ be a $\delta$-error randomized parity decision tree computing
$H^n_k$. Form $A\colon\f\to[0,1]$ so that $A(x)$ is the probability 
$T$ outputs 1 on input $x\in\f$.
Define the polytope
\begin{align*}
\mathcal{P}\defeq\conv
\set{x\mapsto \indicate{}[Bx = c]\mid B\in\field_2^{n\times n}, c\in\f}
\end{align*}
whose vertices are indicator functions for affine subspaces of $\f$.
Given a randomized parity decision tree, 
for each fixing of the randomness,
the set of inputs that end up in
a particular leaf of it is an affine subspace in $\f$. Therefore if
$T$ has at most $s$ leaves, then $A$ is inside $s\mathcal{P}$.
It remains to demonstrate a hyperplane
with normal $H$ so that 
$\iprod{A}{H}>s\iprod{V}{H}$ for any vertex $V$ of the polytope 
$\mathcal{P}$ for $s=\exp \Omega(k\log(k/\delta))$.

Let $\mu_k$ be a distribution on $\f$ obtained as follows. 
Start with the $0$ vector, and flip a coordinate chosen 
uniformly at random with replacement $k$ times. 
Here, flipping a coordinate an even number of times leaves it 
as $0$.
Set $H= \mu_k - (\mu_{k-2}+\mu_{k+2})/(6\delta)$.

First observe that 
$\iprod{A}{\mu_k}>(1-\delta)(1-\binom{k}{2}/n)>1-3\delta/2$
and $\iprod{A}{\mu_{k+2}}+ \iprod{A}{\mu_{k-2}}<3\delta$ so
$\iprod{A}{H}\ge 1/3$ for $\delta\le 1/9$.
Next we would like to upper bound $\iprod{V}{H}$ for an indicator
function $V$ of an affine subspace $\setbuilder{x\in\f}{Bx=c}$.
The key observation is
\begin{align}
\iprod{V}{\mu_k} = \iprod{\indicate{c}}{S^k\indicate{0}}
\label{eq:chi}
\end{align}
where $S$ is a stochastic matrix describing the following transition:
For any $x\in\f$, sample a column $y$ of $B\in\field_2^{n\times n}$ 
uniformly at random and transition to $x+y$. 
Namely, the right hand side of \autoref{eq:chi} describes the following
probability.
We start with the 0 vector in $\f$ and in each time step sample a 
uniform random column $y$ of $B$ and add $y$ to the current state. 
We measure the probability of reaching $c\in\f$ at time step $k$. 
Having observed \autoref{eq:chi}, and that 
$\lnorm{\indicate{0}}{2} = \lnorm{\indicate{c}}{2} = 1$,
the rest of the proof is identical to that of \autoref{thm:klogk}: 
by \autoref{thm:main}, we either have 
\begin{align*}
\iprod{\indicate{c}}{S^{k+2}\indicate{0}}
\iprod{\indicate{c}}{S^{k-2}\indicate{0}}\ge
  \alpha_2 \iprod{\indicate{c}}{S^k\indicate{0}}^2
\end{align*}
or
\begin{align*}
\iprod{\indicate{c}}{S^{k+2}\indicate{0}}
\ge k^{\alpha_1} \iprod{\indicate{c}}{S^k\indicate{0}}^{1+2/k}.
\end{align*}
In either event, we conclude that $\iprod{V}{H}\le 
\nparen{\frac{6\delta}{k^{\alpha_1}}}^{k/2}$.  This completes the proof.
\end{proof}
Note in \autoref{thm:klogk}, we use \autoref{thm:main} with a 
simple and fixed $S$ 
(i.e., the standard random walk on the Hamming cube), 
but with complicated vectors $u,v$ that come from the particular 
communication protocol whose communication cost we would like 
to lower bound. By contrast, in \autoref{thm:paritysize} the 
vectors $u,v$ are simple point masses on states $0$ and $c$ but 
the matrix $S$ is a convolution random walk on the Hamming cube 
that comes from the particular decision tree whose size we lower 
bound.

\subsection{Property testing}
\label{sec:proptest}
In the property testing model, given black box access to 
an otherwise unknown function $f\colon\field^n_2\to\field_2$, 
our goal is to tell apart whether $x\in P$ for some fixed set 
of functions $P$ or $\lone{f-g}\ge \eps2^n$ for any $g\in P$. 
Here, the black box queries are done by providing an input 
$x\in\field_2^n$ to the function and observing $f(x)$.

A function $f\colon\field_2^n\to\field_2$ is
called $k$-linear if $f$ is given by
\begin{align*}
f(x) = \sum_{i\in S}x_i
\end{align*} 
for some $S\subseteq[n]$ of size at most $k$.
By combining our communication complexity lower bound 
\autoref{thm:klogk} with the reduction technique developed 
in \cite{BlaisBM2012} or by combining
our parity decision tree lower bound \autoref{thm:paritysize} 
with a reduction given in \cite{BhrushundiCK2014}, one obtains 
the following.

\begingroup
\def\thecorollary{\ref{cor:propertytest}}
\begin{corollary}[restated]
Any $\delta$-error property testing algorithm for 
$k$-linearity with $\epsilon=1/2$ requires 
$\Omega(k\log (k/\delta))$ queries.
\end{corollary}
\addtocounter{theorem}{-1}
\endgroup
In fact through this, one obtains similar lower bounds to
property testing for $k$-juntas, $k$-term DNFs, size-$k$ formulas,
size-$k$ decision trees, $k$-sparse $\field_2$-polynomials;
see \cite{Blais2009, ChakrabortyGM2011}.
