% !TeX root = heatdiscrete.tex
\section{Monotonicity of 
$t\mapsto m_{2t}^{1/(2t)}$ and 
$t\mapsto m_{2t+1}^{1/(2t+1)}$}
\label{sec:monotonicity}

In this section we prove \autoref{thm:blakley-dixon}
which we restate here 
(with additional equality conditions) 
for the convenience of the reader.
Recall this theorem confirms 
\autoref{conj:blakley-dixon} 
and \autoref{conj:erdos-simonovits}.

\begingroup
\def\thetheorem{\ref{thm:blakley-dixon}}
\begin{theorem}[restated]
Let $S\colon\Omega\times\Omega\to\realspos$ be a 
symmetric matrix with nonnegative entries and 
$u,v\colon\Omega\to\realspos$ be
nonnegative unit vectors. For positive integers 
$k\ge t$ of the same
parity, we have
  \begin{align}
    \iprod{v}{S^ku}^t\ge \iprod{v}{S^tu}^k,
    \label{eq:bd-inthm}
  \end{align}
with equality if and only if $\iprod{v}{S^ku} = 0$ or
$Su =\lambda v$ and $Sv=\lambda u$ for some 
$\lambda\in\realspos$ when $t$ is odd and 
$u=v$ is an eigenvector of $S^2$ when $t$ is even.
\end{theorem}
\addtocounter{theorem}{-1}
\endgroup

We prove \autoref{thm:blakley-dixon} by an 
information theoretic argument.
Define the distributions $\mu \defeq u/\lone{u}$ and 
$\nu \defeq v/\lone{v}$.
Since either side of 
\autoref{eq:bd-inthm} is $kt$-homogeneous in $S$, we may
assume that $S$ is substochastic by scaling as needed. 
Having fixed this normalization, we view 
\autoref{eq:bd-inthm} as a statement about random walks
on $\Omega$ that start from a state sampled according to 
$\mu$ or $\nu$ and evolve according to the transition 
matrix $S$.

\subsection{Reference random walks}
\label{sec:refwalk}
\def\OC{\Omega_\circ}
Let $\OC = \Omega\cup\set{r}$ for some state $r\notin\Omega$ and 
$t$ be a positive integer. Recall that 
$\mu = u/\lone{u}$ and $\nu = v/\lone{v}$.
We start by defining random walks $F^t, B^t$ on $\OC$
that evolve in discrete time steps $-1, 0,1, \ldots, t, t+1$.

The random walk $F^t$ starts at $r$ and transitions to 
a state $x\in\Omega$ with probability $\mu(x)$ at time step $-1$.
In steps $0,1,\ldots, t-1$, the random walk proceeds according to
the transition matrix $S$. At the time step $t$,
each state $x\in\Omega$ transitions to $r$ 
with probability $\nu(x)$ and
transitions to an arbitrary state in 
$\Omega$ with probability
$1-\nu(x)$ 
(say, all of them to the same arbitrary state).
We view $F^t$ as a joint random variable
$F^t = (F^t_{-1},F^t_0,\ldots,F^t_{t+1})$,
where $F^t_i$ is the location of the walk in time step $i$.

The random walk $B^t$ proceeds backwards in time. At time 
step $t+1$ the walk $B^t$ starts at $r$ and transitions
to a state $x\in\Omega$ with probability $\nu(x)$.
In time
steps $t, t-1,\ldots, 1$, the random walk proceeds as prescribed
by $S$.
At time step $0$, each state $x\in\Omega$ transitions
to $r$ with probability $\mu(x)$
and to an arbitrary state in
$\Omega$ with probability $1-\mu(x)$.
Similarly, $B^t$ denotes
the joint random variable 
$B^t = (B^t_{-1},B^t_0,\ldots,B^t_{t+1})$,
where $B^t_i$ is the
location of the walk at time step $i$.

The following facts about $F^t$ and $B^t$ are immediate.
The random variables $F^t_{-1}$ and $B^t_{t+1}$ are fixed
to a single value $r$. The random variables $F^t$, $B^t$ are Markovian,
namely,
$\dist(F^t_i \emid F^t_{i-1},\ldots,F^t_{-1}) 
    = \dist(F^t_i \emid F^t_{i-1})$ and 
$\dist(B^t_{i-1} \emid B^t_{i},\ldots B^t_{t+1})
    = \dist(B^t_{i-1} \emid B^t_{i})$ for 
$i\in \set{0,\ldots, t+1}$.


\subsection{Random walks returning to the origin}
Assume that $\Pr[F^t_{t+1}=r]>0$.
Let $X$ be the walk $F^t$ conditioned on
$F^t_{t+1} = r$.
Note that $X$ is a random variable on the sample space
$\OC^{t+3}$. The next two lemmas explicitly calculate 
the distribution of $X$.

For a matrix $M\colon\Omega\times \Omega\to\realspos$,
functions $f,g\colon\Omega\to\realspos$, and $x,y\in \Omega$
we use the shorthands
\begin{align*}
M(f,y)&\defeq \sum_{x\in\Omega} f(x)M(x,y) = (M^\transpose f)(y)\\
M(x,g)&\defeq \sum_{y\in\Omega} M(x,y)g(y) = (Mg)(x)\\
M(f,g)&\defeq \sum_{x,y\in\Omega} f(x)M(x,y)g(y) 
    = f^\transpose M g,
\end{align*}
where the last expression in each line is understood as
a matrix vector multiplication.
\begin{lemma}
\label{lem:xdist}
Under our assumption $S^t(\mu, \nu) > 0$,
\begin{enumerate}[(i)]
\item we have
  $\Pr\sparen{X_i = x} 
      = \frac{S^i(\mu,x)S^{t-i}(x,\nu)}{S^t(\mu,\nu)}$, and
%
\item if $S^{t-i}(x,\nu)>0$, 
    we have $\Pr[X_{i+1}=y \emid X_{\le i} = x_{\le i}]
    =\frac{S(x_i,y)S^{t-i-1}(y,\nu)}{S^{t-i}(x_i,\nu)}$.
\end{enumerate}
\end{lemma}
\begin{proof}
From the definition of $F^t$ (cf.\ \autoref{sec:refwalk}), 
we have 
  \begin{align}
    \Pr[F^t_i =x] &= S^i(\mu, x)\label{teq:xda}\\
    \Pr[F^t_{t+1}=r\emid F^t_{i}=x] 
        &= S^{t-i}(x,\nu)\label{teq:xdb}\\
    \Pr[F^t_{t+1}=r] &= S^t(\mu,\nu)\label{teq:xdc}\\
    \Pr[F^{t}_{i+1}=y\text{ and }F^t_{t+1}=r\emid 
        F^{t}_i =  x] &= S(x,y)S^{t-i-1}(y,\nu)\label{teq:xde}.
  \end{align}
%TODO(saglam): some more elegant citing mechanism
Using Bayes' rule with \autoref{teq:xda},
\eqref{teq:xdb} and \eqref{teq:xdc} gives (i).
Combining \autoref{teq:xdb}, \eqref{teq:xde}
and the observation that $F^t$ is Markovian
gives (ii).
\end{proof}

With \autoref{lem:xdist} we confirm that the random variable $X=(X_{-1},X_0,\ldots,X_{t+1})$ is Markovian; 
in particular a time inhomogeneous random walk on $\OC$.
Next we observe that the random variable $B^t$ conditioned
on $B^t_{-1}=r$ is precisely $X$ also.

\begin{lemma}
\label{lem:xbackwards}
Under our assumption $S^t(\mu,\nu)>0$,
\begin{enumerate}[(i)]
  \item we have $\dist(X) = \dist(B^t\emid B^t_{-1}=r)$, and
  \item if $S^i(\mu, x)>0$, we have 
      $\Pr[X_{i-1}=y\emid X_{\ge i} = x_{\ge i}] = 
          \frac{S(x_i,y)S^{i-1}(y,\mu)}{S^i(x_i,\mu)}$.
\end{enumerate}
\end{lemma}
\begin{proof}
For any $x\in\OC^{t+3}$ with $x_{t+1} = r$,
  \begin{align*}
    \Pr[X = x] &= \frac{\mu(x_0)\prod_{i=1}^t S(x_{i-1},x_i)\nu(x_t)}
    {S^t(\mu,\nu)}\\
    &=\frac{\nu(x_t)\prod_{i=1}^t S(x_i, x_{i-1}) \mu(x_0)}
        {S^t(\mu,\nu)}
            &&\text{(as $S$ is symmetric)}\\
    &= \frac{\Pr[B^t = x]}{\Pr[B^t_{t+1} = r]}
     = \Pr[B^t = x \emid B^t_{t+1} = r]
        &&\text{(by Bayes' rule).}
  \end{align*}
This proves (i). Given (i), the proof of (ii) is 
the same as \autoref{lem:xdist}(ii).
\end{proof}


\begin{lemma}
\label{lem:xvsf}
We have $\kldiv{X}{F^t} = \kldiv{X}{B^t} = -\log S^t(\mu,\nu)$.
\end{lemma}
\begin{proof}
Recall that $\Pr[F^t_{t+1}=r] = S^t(\mu,\nu)$. 
Since $X$ is obtained from $F^t$ by
conditioning on $F^t_{t+1}=r$, the equality
criteria of \autoref{lem:klcond} are
fulfilled and thus $\kldiv{X}{F^t}=-\log S^t(\mu,\nu)$.
The derivation of $\kldiv{X}{B^t}$ is
identical as per \autoref{lem:xbackwards}(i).
\end{proof}

\subsection{Longer random walks}
\label{sec:zdef}
Let $J$ be an integer valued random variable taking
the values $\set{1,2,\ldots,t}$, each with equal probability. For each fixing
$j$ of $J$ we perform a random walk $Z\emid J=j$ on 
$\OC$ that evolves
in time steps $-1,0,1,\ldots,t,t+1,t+2,t+3$ as follows.

The random walk starts at $r$ and for each time step 
$-1\le i <j$, proceeds according to the
transition kernel $\dist(X_{i+1}\emid X_{i})$. At time step
$j$, the random walk proceeds according to 
$\dist(X_{j-1}\emid X_{j})$ and in time steps 
$j< i \le t+3$ proceeds according to the transition kernel 
$\dist(X_{i-1}\emid X_{i-2})$. We view $Z$ as a joint random
variable $Z=(Z_{-1},Z_0,\ldots,Z_{t+3})$, where $Z_i$ denotes
the location of the random walk at time step $i$.

\begin{lemma}
\label{lem:ydist}
For $-1\le i\le j$, we have 
$\dist(Z_i\emid J=j) = \dist(X_i)$ and for 
$j< i \le t + 3$,
$\dist(Z_i\emid J=j) = \dist(X_{i-2})$.
\end{lemma}
\begin{proof}
This follows from the fact that 
\begin{align*}
\dist(F^t \emid F^t_{t+1} = r) = \dist(X)
    = \dist(B^t \emid B^t_{-1} = r)
\end{align*}
and that $X$ is an actual random walk (i.e., Markovian)
on $\OC$.

To be more explicit, we have $\dist(X_i)=\dist(Z_i)$ for 
$i\le j$ since both $X$ and $Z$ start at $r$ in time step 
$-1$ and evolve according to the transition kernel
$\dist(X_{i+1}\emid X_i)$ for $i=-1,\ldots, j-1$. Since
$\dist(X_j) = \dist(Z_j)$ and $Z$ proceeds according to
$\dist(X_{i-1}\emid X_j)$ at time step $j$, by 
\autoref{lem:xbackwards}, $\dist(X_{j-1}) = \dist(Z_{j+1})$.
%
Finally in time steps $i>j$, we have $\dist(Z_i)=\dist(X_{i-2})$
since $\dist(X_{j-1}) = \dist(Z_{j+1})$ and $Z$ proceeds according
to $\dist(X_{i-1}\emid X_{i-2})$.
\end{proof}

From this we can deduce that $Z$ always
ends up in $r$ at time step $t+3$. We next
argue that if $X$ does not diverge too much from
the reference random walk $F^t$, then $Z$ does 
not diverge too much from $F^{t+2}$.

\begin{lemma}
\label{lem:zdiv}
We have
\begin{align*}
\kldiv{Z\emid J}{F^{t+2}} = \frac{t+2}{t}\kldiv{X}{F^t} 
  &- \frac{1}{t}\nparen{
    \kldiv{X_0}{F^t_0} + 
    \kldiv{X_{t+1}\emid X_{t}}{F^t_{t+1}\emid F^t_t}
  }\\
  &- \frac{1}{t}\nparen{
    \kldiv{X_t}{B^t_t} + 
    \kldiv{X_{-1}\emid X_{0}}{B^t_{-1}\emid B^t_0}
  }.
\end{align*}
\end{lemma}
\begin{proof}
For a fixing $j$ of $J$, we have 
\begin{align*}
\kldiv{Z\emid J=j}{F^{t+2}} 
    &= \sum_{i=-1}^{j-1}\kldiv{X_{i+1} \emid X_{i}}
        {F^{t+2}_{i+1} \emid F^{t+2}_{i}}
    + \kldiv{X_{j-1}\emid X_j}{F^{t+2}_{j+1}
                    \emid F^{t+2}_{j}}\\
    &\quad\quad
        + \sum_{i=j+1}^{t+2}\kldiv{X_{i-1}\emid X_{i-2}}
      {F^{t+2}_{i+1}\emid F^{t+2}_{i}},
\end{align*}
where we have used the chain rule for divergence (cf.\ 
\autoref{lem:klchain}), the fact that $Z\emid J=j$ and 
$F^{t+2}$ are Markovian and \autoref{lem:ydist}. 
Recalling that $F^t$ and $B^t$ evolve according 
to $S$ in time steps $0,1,\ldots,t-1$, and 
$\dist(F^t_{t+1}\emid F^t_t)
    =\dist(F^{t+2}_{t+3}\emid F^{t+2}_{t+2})$,
we write
\begin{align*}
\kldiv{Z\emid J=j}{F^{t+2}}
  &= \sum_{i=-1}^t \kldiv{X_{i+1}\emid X_i}{F^t_{i+1}\emid F^t_i}\\
     &\qquad\qquad
     + \kldiv{X_{j-1}\emid X_j}{B^t_{j-1}\emid B^t_{j}}
     + \kldiv{X_j\emid X_{j-1}}{F^t_{j}\emid F^t_{j-1}}\\
  &= \kldiv{X}{F^t} 
    + \kldiv{X_{j-1}\emid X_j}{B^t_{j-1}\emid B^t_{j}}
    + \kldiv{X_j\emid X_{j-1}}{F^t_{j}\emid F^t_{j-1}}
\end{align*}
again by the chain rule for divergence 
(\autoref{lem:klchain}) and the
fact that $X$ and $F^t$ are Markovian. Now taking 
an expectation over all $j\in \supp(J)$, we have
\begin{align*}
\kldiv{Z\emid J}{F^{t+2}} 
  &= \frac{1}{t}\sum_j\kldiv{Z\emid J=j}{F^{t+2}}\\
  &= \kldiv{X}{F^t} 
    + \frac{1}{t}\sum\nparen{\kldiv{X_{j-1}\emid X_j}
                            {B^t_{j-1}\emid B^t_{j}}
    + \kldiv{X_j\emid X_{j-1}}{F^t_{j}\emid F^t_{j-1}}}\\
  &= \kldiv{X}{F^t}
  + \frac{\kldiv{X}{B^t} 
      - \kldiv{X_t}{B^t_t}
      - \kldiv{X_{-1}\emid X_0}{B^t_{-1}\emid B^t_0}
    }{t}\\
  &\qquad\qquad\qquad\;+ \frac{\kldiv{X}{F^t} 
      - \kldiv{X_0}{F^t_0}
      - \kldiv{X_{t+1}\emid X_t}{F^t_{t+1}\emid F^t_t}
    }{t}.
\end{align*}
Since $\kldiv{X}{B^t} = \kldiv{X}{F^t}$ by 
\autoref{lem:xvsf}, collecting the $\kldiv{X}{F^t}$
terms we finish the proof.
\end{proof}

Finally, we lower bound the negative terms in the statement
of \autoref{lem:zdiv}.
\begin{lemma}
\label{lem:negterms}
We have
\begin{align}
\kldiv{X_0}{F^t_0}
    + \kldiv{X_{-1}\emid X_{0}}{B^t_{-1}\emid B^t_0}
    &\ge \ryent{2}{\mu} \defeq -\log \lnorm{\mu}{2}^2,
        \text{ and} \label{teq:negterm1}\\
\kldiv{X_t}{B^t_t}
    + \kldiv{X_{t+1}\emid X_{t}}{F^t_{t+1}\emid F^t_t}
    &\ge \ryent{2}{\nu} \defeq -\log \lnorm{\nu}{2}^2,
        \label{teq:negterm2}
\end{align}
where $\ryent{2}{\cdot}$ denotes the 
second order Rényi entropy. 
%If the inequalities hold
%with equality simultaneously, then $\mu = \nu$ and 
%$\mu$ is an eigenvector of $S$.
\end{lemma}
\begin{proof}
We only prove the first inequality as the
second one is symmetric. By \autoref{lem:xdist},
we have
\begin{align*}
\kldiv{X_0}{F^t_0} &= 
  \sum_{x\in \Omega}
      \frac{\mu(x)S^{t}(x,\nu)}{S^t(\mu,\nu)}
      \log \frac{S^{t}(x,\nu)}{S^t(\mu,\nu)}
          \quad\text{ and}\\
\kldiv{X_{-1}\emid X_{0}}{B^t_{-1}\emid B^t_0} &=
  \sum_{x\in \Omega}
      \frac{\mu(x)S^{t}(x,\nu)}{S^t(\mu,\nu)}
      \log\frac{1}{\mu(x)}.
\end{align*}
Let $\Psi=\supp(X_0)$. By adding the two terms we get
\begin{align}
\kldiv{X_0}{F^t_0}+
\kldiv{X_{-1}\emid X_{0}}{B^t_{-1}\emid B^t_0} &=
  -\sum_{x\in\Psi}
      \frac{\mu(x)S^{t}(x,\nu)}{S^t(\mu,\nu)}
      \log\frac{\mu(x)S^t(\mu,\nu)}{S^{t}(x,\nu)}
          \nonumber\\
  &\ge -\log \sum_{x\in\Psi}\frac%
    {\mu(x)^2 S^{t}(x,\nu)S^t(\mu,\nu)}
      {S^t(\mu,\nu)S^{t}(x,\nu)}
          \label{eq:norm-u-jens}\\
  &=   -\log \sum_{x\in\Psi} \mu(x)^2
      \nonumber\\
  &\ge -\log \sum_{x\in\Omega} \mu(x)^2\;,
      \label{eq:psivsomega}
\end{align}
where the first inequality is by concavity of 
$z\mapsto\log z$
and the second inequality is true as the 
summands are nonnegative.
\end{proof}

\subsection{Combining the inequalities}
\label{sec:bd-final}
\begin{proof}[Proof of \autoref{thm:blakley-dixon}]
Note that $Z_{-1}$ is fixed to $r$ by definition 
(cf.\ \autoref{sec:zdef}) and $Z_{t+3}$ is fixed to $r$
by \autoref{lem:ydist}. Therefore by 
\autoref{lem:klcond} we have
\begin{align}
-\log S^{t+2}(\mu,\nu)
  &\le \kldiv{Z}{F^{t+2}} \label{eq:kvk2} \\
  &= \kldiv{Z\emid J}{F^{t+2}} - \muti{J}{Z}
      \label{eq:ztozcondj}\\
  &\le \kldiv{Z\emid J}{F^{t+2}}
      \label{eq:mutrelax}\\
  &\le \frac{t+2}{t}\kldiv{X}{F^t} 
      + \frac{\log\lnorm{\mu}{2}^2 + \log\lnorm{\nu}{2}^2}
            {t}.
        \nonumber
\intertext{%
Here \autoref{eq:ztozcondj} follows from the chain rule 
for the divergence (\autoref{lem:klchain}) 
and the definition of mutual information 
(cf.\ \autoref{eq:muti-def}),
\autoref{eq:mutrelax} follows from
the nonnegativity of mutual information and
the last line follows from 
\autoref{lem:zdiv} and \autoref{lem:negterms}.
Plugging in $\kldiv{X}{F^t}=-\log S^t(\mu,\nu)$, provided by 
\autoref{lem:xvsf}, we obtain}
  -\log S^{t+2}(\mu,\nu)
    &\le -\frac{t+2}{t}\log S^t(\mu,\nu) 
         + \frac{\log\lnorm{\mu}{2}^2 + \log\lnorm{\nu}{2}^2}
            {t}.
        \nonumber
\end{align}
Arranging, we get
\begin{align*}
\lnorm{\mu}{2}^2 \lnorm{\nu}{2}^2\iprod{\nu}{S^{t+2}\mu}^t
  \ge
    \iprod{\nu}{S^t\mu}^{t+2}
\end{align*}
and substituting $\mu = u/\lone{u}$, 
$\nu = v/\lone{v}$, and recalling that $u,v$ 
are unit vectors, we obtain
\begin{align}
\iprod{v}{S^{t+2}u}^t 
&\ge \iprod{v}{S^t u}^{t+2}, \text{ i.e.,} \nonumber\\
m_{t+2} &\ge m_t^{1+2/t}\label{eq:absclaim1}.
\end{align}
By applying this inequality iteratively, we get
$\iprod{v}{S^{k}u}^t\ge\iprod{v}{S^t u}^{k}$ or
written differently
$m_k^{1/k}\ge m_t^{1/t}$
as long as $k>t$ and $k,t$ have the same parity.

Next we characterize the equality conditions of 
\autoref{eq:absclaim1}. Let us verify the
`if' direction of the statement. 
Clearly if $\iprod{v}{S^ku}=0$ then we have $\iprod{v}{S^ku} = \iprod{v}{S^tu}$ by the
first part of the theorem and the fact that 
$\iprod{v}{S^tu}\ge 0$. If $S^2u = \lambda^2 u$, 
then $m_{2t} = \lambda^{2t}$ and if $Su=\lambda v$ and $Sv=\lambda u$, then $m_{2t+1}=\lambda^{2t+1}$, therefore
in both $t$ even and $t$ odd cases 
the inequality holds with equality.

Conversely, if 
$0\neq \iprod{v}{S^{k+2}u}=\iprod{v}{S^ku}$, then the
inequalities
\eqref{eq:norm-u-jens}, 
\eqref{eq:psivsomega},
\eqref{eq:kvk2},
 and
\eqref{eq:ztozcondj} must hold with equality.
Combining the assumption that \autoref{eq:psivsomega}
and \eqref{eq:norm-u-jens} 
hold with equality with the strict concavity of
$z\mapsto \log z$ and Jensen's lemma, we get that 
$S^t\nu = \lambda_1\mu + \sigma_1$ for some $\lambda_1\le 1$ and $\sigma_1\in\reals^{\Omega}_{+}$ satisfying 
$\supp(\sigma_1)\cap\supp(\mu)=\emptyset$. This also means that $\Pr[X_0 = x] = \mu(x)^2/\lnorm{\mu}{2}^2$ for $x\in \Omega$ and a similar and symmetrical argument shows that
$\Pr[X_t = x] = \nu(x)^2/\lnorm{\nu}{2}^2$ for 
$x\in\Omega$.
Assuming \autoref{eq:ztozcondj} holds 
with equality leads to $\muti{Z}{J}=0$, which in turn
shows that $\dist(X_i)=\dist(X_{i+2})$ for 
$i=0,\ldots,t-2$. 
Let 
$X^{k+2} \defeq \nparen{F^{t+2}\emid F^{t+2}_{t+3}=r}$.
From our assumption $0\neq \iprod{v}{S^{k+2}u}=\iprod{v}{S^ku}$ we conclude that $\dist(Z)=\dist(X^{t+2})$
as $X^{t+2}$ is the minimally divergent distribution from $F^{t+2}$ among distributions on walks ending at state $r$. Since 
$\dist(X^{t+2}_2)=\dist(X^{t+2}_0)=\dist(X_0)$, and $S^t\nu=\lambda_1\mu+\sigma_1$ we get that 
$S^2\mu = \lambda_2 \mu + \sigma_2$  for some $\lambda_2\le 1$ and $\sigma_2\in\reals^{\Omega}_{+}$ satisfying 
$\supp(\sigma_2)\cap\supp(\mu)=\emptyset$. Now we will show that it must be that $\sigma_2=0$. Suppose for sake of contradiction that $\sigma_2(z)>0$ for some $z\notin \supp(\mu)$. There exists $x,y\in\Omega$ so that $\mu(x)S(x,y)S(y,z)>0$. If $y\in\supp(X_1)$ then adding to a walk $w\in \supp(X)$ with $w_1=y$ the loop $(y,z)(z,y)$
we obtain a length $t+2$ walk which is not in the support of $X^{t+2}$ as $\dist(X^{t+2}_2)=\dist(X_2)=\dist(X_0)$,
which contradicts the fact that 
$X^{t+2}$ is defined as $F^{t+2}\emid F^{t+2}_{t+3}=r$.
If on the other hand $y\notin\supp(X_1)$, 
adding the loop $(x,y)(y,x)$ to a walk with $w_0=x$ 
leads to a walk which is not in the support of $X^{t+2}$, 
which is a contradiction. Having established $S^2\mu = \lambda \mu$, we complete the proof for even $t$
by recalling that $\dist(X_0)=\dist(X_t)$ therefore $\mu=\nu$. For $t$ odd, the last argument shows that
 $S\mu = \lambda_3\nu + \sigma_3$ for some $\lambda_3\le 1$ and $\sigma_3\in\reals^{\Omega}_{+}$ satisfying 
$\supp(\sigma_3)\cap\supp(\mu)=\emptyset$. 
It remains to show that $\sigma_3=0$ 
by using the assumption that $\dist(Z)=\dist(X^{t+2})$. Suppose $\sigma_3(y)>0$ for some $y\notin\supp(\nu)$. There exists $x\in\Omega$ such that $\mu(x)S(x,y)>0$. Adding the loop $(x,y)(y,x)$ to
a walk $w$ with $w_{t-1}=x$ leads to a length $t+2$ walk
which is not in the support of $X^{t+2}$. A symmetrical
argument shows that $\lambda'\mu = S\nu$. This completes the proof.
\end{proof}